### 2019-12-12 tim-lucas-model-ensembles-SI.tex
Add SI

\documentclass[review]{elsarticle}

\usepackage{geometry}
 \geometry{
 a4paper,
 left=30mm,
 right=30mm,
 top=10mm,
 bottom=10mm
 }

\usepackage{graphicx}
\usepackage{amsmath}

\usepackage{subcaption}

\usepackage{amssymb}
\usepackage{array}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

%\usepackage{numcompress}\bibliographystyle{model3-num-names}

\usepackage{natbib}
%\usepackage[numbers, sort&compress]{natbib}
%\setcitestyle{author[oxford]year}
\bibliographystyle{elsarticle-num}

%\geometry{verbose,tmargin=30mm,bmargin=25mm,lmargin=25mm,rmargin=25mm}
%\input{figures.tex}
%\pagestyle{empty}
\begin{document}
%\templatefigures{}

% Model ensembles with different response variables for base and meta models:

% Spatial and Spatio-temporal Epidemiology
% Special issue: Disease Surveillance and Infectious disease Modeling

\begin{frontmatter}

\title{Supporting information: Improving disaggregation models of malaria incidence by ensembling non-linear models of prevalence}

\author[oxford]{Tim C. D. Lucas\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{timcdlucas@gmail.com}
\author[oxford]{Anita K. Nandi}
\author[oxford]{Suzanne H. Keddie}
%\author[oxford]{Katherine A. Twohig}
\author[oxford]{Elisabeth G. Chestnutt}
\author[oxford]{Rosalind E. Howes}
%\author[oxford]{Michele Nguyen}
\author[oxford]{Susan F. Rumisha}
\author[oxford]{Rohan Arambepola}
\author[oxford,idm]{Amelia Bertozzi-Villa}
\author[oxford]{Andre Python}
\author[oxford]{Tasmin L. Symons}
\author[oxford]{Justin J. Millar}
\author[oxford]{Punam Amratia}
\author[oxford]{Penelope Hancock}
\author[oxford]{Katherine E. Battle}
\author[oxford]{Ewan Cameron}
\author[oxford,telethon,curtin]{Peter W. Gething}
\author[oxford]{Daniel J. Weiss}

\address[oxford]{Malaria Atlas Project, Big Data Institute, University of Oxford, Oxford, UK}
\address[idm]{Institute for Disease Modeling, Bellevue, WA, USA}
\address[telethon]{Telethon Kids Institute, Perth Children’s Hospital, Perth, Australia}
\address[curtin]{Curtin University, Perth, Australia}

\end{frontmatter}

\clearpage
\tableofcontents

\clearpage
\section{Global dataset Machine Learning}

\clearpage
\subsection{Out-of-sample scatter plots}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnet_obspred_global.png}
\caption{
  Scatter plot of predictions and held out observed data for the neural network trained on the global dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enet_obspred_global.png}
\caption{
  Scatter plot of predictions and held out observed data for the elastic net trained on the global dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppr_obspred_global.png}
\caption{
  Scatter plot of predictions and held out observed data for the PPR trained on the global dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ranger_obspred_global.png}
\caption{
  Scatter plot of predictions and held out observed data for the Random Forest trained on the global dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/gbm_obspred_global.png}
\caption{
  Scatter plot of predictions and held out observed data for the GBM trained on the global dataset.
}

\end{figure}

\clearpage
\subsection{Hyperparameter optimisation}

As ranger and GBM were tuned with random hyperparameter search, the plots become difficult and are not included.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnetopt_global.png}
\caption{
  Optimisation for neural network hyperparameters trained on the global dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enetopt_global.png}
\caption{
  Optimisation for elastic net hyperparameters trained on the global dataset.
}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppropt.png}
\caption{
  Optimisation for PPR hyperparameters trained on the global dataset.
}

\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Colombia (South America) prevalence dataset Machine Learning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Predictions}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/SA_all_ml.png}
\caption{
  Predictions from machine learning models trained on South American prevalence data.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/COL_all_globalml.png}
\caption{
  Predictions over Colombia from machine learning models trained on global prevalence data.
}

\end{figure}

\clearpage
\subsection{Out-of-sample scatter plots}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnet_obspred_sa.png}
\caption{
  Scatter plot of predictions and held out observed data for the neural network trained on the South America dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enet_obspred_sa.png}
\caption{
  Scatter plot of predictions and held out observed data for the elastic net trained on the South America dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppr_obspred_sa.png}
\caption{
  Scatter plot of predictions and held out observed data for the PPR trained on the South America dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ranger_obspred_sa.png}
\caption{
  Scatter plot of predictions and held out observed data for the Random Forest trained on the South America dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/gbm_obspred_sa.png}
\caption{
  Scatter plot of predictions and held out observed data for the GBM trained on the South America dataset.
}

\end{figure}

\clearpage
\subsection{Hyperparameter optimisation}

As ranger and GBM were tuned with random hyperparameter search, the plots become difficult and are not included.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnetopt_sa.png}
\caption{
  Optimisation for neural network hyperparameters trained on the South America dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enetopt_sa.png}
\caption{
  Optimisation for elastic net hyperparameters trained on the South America dataset.
}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppropt_sa.png}
\caption{
  Optimisation for PPR hyperparameters trained on the South America dataset.
}

\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Indonesia dataset Machine Learning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Predictions}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/idn_all_ml.png}
\caption{
  Predictions from machine learning models trained on Indonesian prevalence data.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/idn_all_globalml.png}
\caption{
  Predictions over Indonesia from machine learning models trained on global prevalence data.
}

\end{figure}

\clearpage
\subsection{Out-of-sample scatter plots}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnet_obspred_idn.png}
\caption{
  Scatter plot of predictions and held out observed data for the neural network trained on the Indonesia dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enet_obspred_idn.png}
\caption{
  Scatter plot of predictions and held out observed data for the elastic net trained on the Indonesia dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppr_obspred_idn.png}
\caption{
  Scatter plot of predictions and held out observed data for the PPR trained on the Indonesia dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ranger_obspred_idn.png}
\caption{
  Scatter plot of predictions and held out observed data for the Random Forest trained on the Indonesia dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/xgboost_obspred_idn.png}
\caption{
  Scatter plot of predictions and held out observed data for the GBM trained on the Indonesia dataset.
}

\end{figure}

\clearpage
\subsection{Hyperparameter optimisation}

As ranger and GBM were tuned with random hyperparameter search, the plots become difficult and are not included.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enetopt_idn.png}
\caption{
  Optimisation for elastic net hyperparameters trained on the Indonesia dataset.
}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppropt_idn.png}
\caption{
  Optimisation for PPR hyperparameters trained on the Indonesia dataset.
}

\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Madagascar dataset Machine Learning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Predictions}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/MDG_all_ml.png}
\caption{
  Predictions from machine learning models trained on Malagasy prevalence data.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/MDG_all_globalml.png}
\caption{
  Predictions over Madagascar from machine learning models trained on global prevalence data.
}

\end{figure}

\clearpage
\subsection{Out-of-MDGmple scatter plots}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnet_obspred_mdg.png}
\caption{
  Scatter plot of predictions and held out observed data for the neural network trained on the Madagascar dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enet_obspred_mdg.png}
\caption{
  Scatter plot of predictions and held out observed data for the elastic net trained on the Madagascar dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppr_obspred_mdg.png}
\caption{
  Scatter plot of predictions and held out observed data for the PPR trained on the Madagascar dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ranger_obspred_mdg.png}
\caption{
  Scatter plot of predictions and held out observed data for the Random Forest trained on the Madagascar dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/xgboost_obspred_mdg.png}
\caption{
  Scatter plot of predictions and held out observed data for the GBM trained on the Madagascar dataset.
}

\end{figure}

\clearpage
\subsection{Hyperparameter optimiMDGtion}

As ranger and GBM were tuned with random hyperparameter search, the plots become difficult and are not included.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enetopt_mdg.png}
\caption{
  OptimiMDGtion for elastic net hyperparameters trained on the Madagascar dataset.
}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppropt_mdg.png}
\caption{
  OptimiMDGtion for PPR hyperparameters trained on the Madagascar dataset.
}

\end{figure}

\clearpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Senegal dataset Machine Learning}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Predictions}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/sen_all_ml.png}
\caption{
  Predictions from machine learning models trained on Senegal prevalence data.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=1\textwidth]{figs/SI/sen_all_globalml.png}
\caption{
  Predictions from machine learning models trained on Senegal prevalence data.
}

\end{figure}

\clearpage
\subsection{Out-of-sample scatter plots}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/nnet_obspred_sen.png}
\caption{
  Scatter plot of predictions and held out observed data for the neural network trained on the Senegal dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enet_obspred_sen.png}
\caption{
  Scatter plot of predictions and held out observed data for the elastic net trained on the Senegal dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppr_obspred_sen.png}
\caption{
  Scatter plot of predictions and held out observed data for the PPR trained on the Senegal dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ranger_obspred_sen.png}
\caption{
  Scatter plot of predictions and held out observed data for the Random Forest trained on the Senegal dataset.
}

\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/xgboost_obspred_sen.png}
\caption{
  Scatter plot of predictions and held out observed data for the GBM trained on the Senegal dataset.
}

\end{figure}

\clearpage
\subsection{Hyperparameter optimisation}

As ranger and GBM were tuned with random hyperparameter search, the plots become difficult and are not included.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/enetopt_sen.png}
\caption{
  Optimisation for elastic net hyperparameters trained on the Senegal dataset.
}
\end{figure}

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/SI/ppropt_sen.png}
\caption{
  Optimisation for PPR hyperparameters trained on the Senegal dataset.
}

\end{figure}

%\section*{Citations}
%\bibliography{Malaria} 
%\bibliographystyle{unsrt}
%\bibliographystyle{apalike}

\end{document}